---
title: "Basic Stat Lab"
format: html
editor: visual
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.height = 3.5, 
                      fig.width  = 6,
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r lib}
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(Sleuth3)
library(broom)
```

# Basic Statistics Classwork/Lab

1)  In Hong Kong, human male height is approximately normally distributed with mean 171.5 cm and standard deviation 5.5 cm. (Use and show R code to produce answers for a -- e) You can use any method for part f.\
    a) What proportion of the Hong Kong population is between 170 cm and 180 cm?\
```{r 1a}
aa1 <- pnorm(180, mean=171.5, sd=5.5)
aa2 <- pnorm(170, mean=171.5, sd=5.5)
aa1-aa2
```
    b) What proportion of the Hong Kong population is less that 165 cm ?\
```{r 1b}
pnorm(165, mean=171.5, sd=5.5)
```
    c) What proportion of the Hong Kong population is greater than 175 ?\
```{r 1c}
aa3 <- pnorm(175, mean=171.5, sd=5.5)
1-aa3
```
    d) The proportion .25 is less than what male height value ?\
```{r 1d}
qnorm(p = 0.25, mean = 171.5, sd = 5.5)
```
    e) The proportion .85 is greater than what male height value ?\
```{r 1e}
qnorm(p = 0.849999, mean = 171.5, sd = 5.5)
```
    f) The proportion of .68 is between what two male height values ?\
```{r 1f}
qnorm(p = 0.679, mean = 171.5, sd = 5.5)
qnorm(p = 0.681, mean = 171.5, sd = 5.5)
```

2)  Use and show R code to determine the height of the standard normal curve at a Z value of 2.5.
```{r 2}
dnorm(x = 2.5, mean = 0, sd = 1)
```
Because the Z-Score is z = (value - mean)/(stdv) and the mean = 0 & stdv = 1 for a normal curve, we can just plug in the z-score being asked for bc z = (value - mean)/(stdv) = (value - 0)/(1) = value.

3)  Use and show R code to determine the height of the standard normal curve at a Z value of .4.
```{r 3}
dnorm(x = .4, mean = 0, sd = 1)
```
4)  Use and show R code to determine the height of a t distribution curve for a t value of 1.5 with 3 degrees of freedom.
```{r 4}
dt(1.5, 3)
```
5)  Use and show R code to determine the height of a t distribution curve for a t value of -2 with 2 degrees of freedom.
```{r 5}
dt(-2, 2)
```
6)  For a t distribution with 3 degrees of freedom, use and show R code that will find the proportion less than 2.5.
```{r 6}
pt(2.5,3)
```
7)  For a t distribution with 1 degree of freedom, use and show R code that will find the proportion that is greater than 1.75.
```{r 7}
aa4 <- pt(1.75,1)
1-aa4
```
8)  For a t distribution with 2 degrees of freedom, use and show R code that will find the value immediately above a proportion of .355.
```{r 8}
qt(.3551,2)
```
9)  An educator believes that new directed reading activities in the classroom will help elementary school pupils improve some aspects of their ability. She arranges for a third -grade class of 23 students to take part in these activities for an eight-week period. A control classroom of 23 third graders follows the same curriculum without the activities. At the end of the eight weeks, all students are given a Degree of Reading Power (DRP) test, which measures the aspects of reading ability that the treatment is designed to improve. The sample data performance results are provided below;

-   Treatment Group: 24,61,59,46,43,44,52,43,58,67,62,57,71,49,54,43,53,57,49,56,33,74,70

-   Control Group: 42,33,46,37,43,41,10,42,55,19,17,55,26,54,60,28,62,20,53,48,37,85,42

-   Design and execute a two sample t test.\
a) State the appropriate null and alternative hypotheses.
H0: Treatment mean = Control mean
Ha: Treatment mean > Control mean

b)  Use and show R code to produce the p value and the confidence interval 
```{r 9b}
df_9b <- tibble(Treatment= c(24,61,59,46,43,44,52,43,58,67,62,57,71,49,54,43,53,57,49,56,33,74,70), Control=c(42,33,46,37,43,41,10,42,55,19,17,55,26,54,60,28,62,20,53,48,37,85,42))
df_9b
t.test(df_9b$Treatment,df_9b$Control, alternative = "greater")
```
c) Determine if you should reject or fail to reject the null hypothesis using the p value and the         confidence interval.

As p=0.005271<0.05 we conclude that there is sufficient statistical evidence to reject the null hypothesis.